---
title: "quarto"
editor_options: 
  chunk_output_type: console
bibliography: fid_article.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(chromatographR)
library(rdryad)
library(vegan)
library(ggplot2)
```

### Introduction

While the basic workflow is similar for analyzing different types of chromatographic data with `chromatographR`, there are a few important differences when analyzing GC-FID data (or other 1-dimensional data). This vignette highlights some important considerations when using chromatographR to analyze GC-FID data (or other 2D chromatography data).

To demonstrate the use of `chromatographR` for the analysis GC-FID data, we will analyze a dataset on cuticular hydrocarbons (CHCs) composition in four species of paper wasps (*Polistes spp.*) collected by Dr. Andrew Legan [@legan2022; @legan2022a]. The dataset includes CHCs collected from *Polistes dominula* (european paper wasp), *Polistes exclamans* (common paper wasp), *P. fuscatus* (northern paper wasp), and *P. metricus* (metric paper wasp). As in other social insects, cuticular hydrocarbons serve as chemical cues mediating a number of social behaviors in paper wasps [@legan2021], including nestmate recognition [@gamboa1986; @gamboa1996; @bruschini2011], establishment of social hierarchies [@jandt2014] and mate choice [@reed1990].

```{r download_data}
# load chromatograms from Legan et al 2022
files <- rdryad::dryad_download("10.5061/dryad.wpzgmsbr8")[[1]]
chrom_paths <- grep("README|METADATA|ANNOTATED", files, invert = TRUE, value = TRUE)
dat <- read_chroms(chrom_paths, format_in = "shimadzu_fid")

# load metadata
path_to_metadata <- grep("METADATA", files, value = TRUE)
meta <- read.csv(path_to_metadata, sep = "\t")
```
